---
title: "DATA 608 Final Project"
author: "Vinicio Haro"
date: "May 7, 2019"
output: html_document
---

This document is to put together an outline and explore some of the potential data sources for the final DATA 608 project. 

The goal is to visually explore the relationship between low income neighborhoods and poverty within New York City. I want to be able to visually show these elements using a combination of tatic plots and and shiny apps.

NYC.gov has a very handy export tool that allows us to easily extract the data we need. We will pull poverty data by neighborhood and obesity data by boro. Povery data will need to be rolled up in order to match up with the obesity data since they are both two different types of granularity. 

http://a816-dohbesp.nyc.gov/IndicatorPublic/VisualizationData.aspx?id=103,4466a0,109,Summarize

http://a816-dohbesp.nyc.gov/IndicatorPublic/VisualizationData.aspx?id=2063,4466a0,113,Summarize

https://www.census.gov/topics/income-poverty/income.html

This data will be stored in github and pulled directly from github. 
```{r message=FALSE, warning=FALSE, include=FALSE}
library("tmap")
library("tmaptools")
library("sf")
library("leaflet")
library(ggplot2)
library(sp)
library(maps)
```


## Read in the data sources
```{r message=FALSE, warning=FALSE}
library(readr)
poverty.df <- read_csv("https://raw.githubusercontent.com/vindication09/DATA-608/master/Poverty%20(2).csv")

obesity.df <- read_csv("https://raw.githubusercontent.com/vindication09/DATA-608/master/Obese%20Adults%20(2).csv")

national<-read_csv("https://raw.githubusercontent.com/vindication09/DATA-608/master/National_Obesity_By_State.csv")

income<-read_csv("https://raw.githubusercontent.com/vindication09/DATA-608/master/2015incomenew.csv")

state_income<-read_csv("https://raw.githubusercontent.com/vindication09/DATA-608/master/statelevel_medianincome.csv")

state_obesity<-read_csv("https://raw.githubusercontent.com/vindication09/DATA-608/master/cdc_obesity_survey.csv")

head(poverty.df);head(obesity.df);head(national.df);head(income);head(state_income);head(state_obesity)
```

Lets examine poverty on the national level
```{r message=FALSE, warning=FALSE}
library(usmap)
library(ggplot2)
library(ggrepel)
library(sqldf)

nationalb<-sqldf("select 
                 Name as 'state',
Obesity
from national")

usmap::plot_usmap(regions = "counties") + 
  labs(title = "US Counties", subtitle = "This is a blank map of the counties of the United States.") + 
  theme(panel.background = element_rect(colour = "black", fill = "lightblue"))

summary(nationalb)
```

##Mapping the Obesity Rate in 2015

```{r warning=FALSE, message=FALSE}


usmap::plot_usmap(data = nationalb, values = "Obesity", lines = "black") + 
  scale_fill_continuous(name = "Percent",low = "white", high = "red", label = scales::comma) + 
  theme(legend.position = "right")+ 
  labs(title = "United States Obesity Distribution", subtitle = "The percent of the state population that is considered obese from the 2015 CDC BRFSS Survey.")
```

source: http://data-lakecountyil.opendata.arcgis.com/datasets/3e0c1eb04e5c48b3be9040b0589d3ccf_8

## Mapping the Household income by State for (2015)
```{r r warning=FALSE, message=FALSE}
incomeb<-sqldf("select 
              State as state, 
               income
               from income")


usmap::plot_usmap(data = incomeb, values = "income", lines = "red") + 
  scale_fill_continuous(name = "HHI",low = "white", high = "green", label = scales::comma) + 
  theme(legend.position = "right")+ 
  labs(title = "United States Median Household Income", subtitle = "2015 Census")


```


https://cran.r-project.org/web/packages/usmap/vignettes/mapping.html



The above plots only give us a snapshot of income and obesity during a given year. We would need a group of small multiples, however we can better display this in a shiny app allowing the user to control the state. 

```{r warning=FALSE, message=FALSE}
#extract the state level data 
stateonly_income<-sqldf("select 
                    * 
                    from state_income 
                    where state not in ('United States')")

national_income<-sqldf("select
                       *
                       from state_income
                       where state='United States'")

#bring these two data sources together 
joina<-sqldf("select 
             a.year, 
             a.state,
             a.income,
             b.income as national_avg_income
             from stateonly_income a
             join national_income b 
             on (a.year=b.year)")
```


##Shiny Household income per state vs national over time 
```{r}
library(shiny)
library(plotly)

ui <- fluidPage(
  sidebarPanel(
    selectInput(inputId = "state", label = "State Selected:",
                choices = levels(as.factor(joina$state))
    ),
 
    helpText("Median Household Income per State vs National Median Household Income"),
    width = "auto"
  ),
  plotOutput("plot2")
)

server<- function(input, output) {
  
  output$plot2 <-renderPlot({
    
    SelectedState <- input$state
    
    ggplot(joina[joina$state == SelectedState,]) +
      geom_bar(aes(x = year, weight = income), fill = "steelblue") +
      labs(x="Year", y = "Median Household Income", 
           title = "U.S Median Household Income",
           subtitle = paste("U.S Census 2013-2017")) +
      geom_line(aes(x = year, y = national_avg_income, linetype = "National Median HHI"), col = "red", lwd = 1) +
      scale_linetype(name = NULL) +
      theme_bw()
  })
}

shinyApp(ui = ui, server = server)

```

##Shiny obesity Rate per State vs National Obesity Average 
data: https://chronicdata.cdc.gov/Nutrition-Physical-Activity-and-Obesity/Nutrition-Physical-Activity-and-Obesity-Behavioral/hn4x-zwk7

https://nccd.cdc.gov/dnpao_dtm/rdPage.aspx?rdReport=DNPAO_DTM.ExploreByTopic&islClass=OWS&islTopic=OWS1&go=GO

This dataset includes data on adult's diet, physical activity, and weight status from Behavioral Risk Factor Surveillance System. This data is used for DNPAO's Data, Trends, and Maps database, which provides national and state specific data on obesity, nutrition, physical activity, and breastfeeding.

Lets look at descriptive statistics after we seperate the average from the state level 
```{r}
#extract the state level data 
stateonly_obesity<-sqldf("select 
                    * 
                    from state_obesity
                    where state not in ('United States')")

national_obesity<-sqldf("select
                       *
                       from state_obesity
                       where state='United States'")

#bring these two data sources together 
joinb<-sqldf("select 
             a.year, 
             a.state,
             a.Data_Value as state_value,
             b.Data_Value as national_value
             from stateonly_obesity a
             join national_obesity b 
             on (a.year=b.year)")
```

App
```{r}
ui <- fluidPage(
  sidebarPanel(
    selectInput(inputId = "state", label = "State Selected:",
                choices = levels(as.factor(joinb$state))
    ),
 
    helpText("Percent of Adults Classified as Obese"),
    width = "auto"
  ),
  plotOutput("plot3")
)

server<- function(input, output) {
  
  output$plot3 <-renderPlot({
    
    SelectedState <- input$state
    
    ggplot(joinb[joinb$state == SelectedState,]) +
      geom_bar(aes(x = year, weight = state_value), fill = "steelblue") +
      labs(x="Year", y = "Percent of Adults Classified as obese", 
           title = "U.S CDC Obesity Survey",
           subtitle = paste("CDC 2013-2017")) +
      geom_line(aes(x = year, y = national_value, linetype = "National Average"), col = "red", lwd = 1) +
      scale_linetype(name = NULL) +
      theme_bw()
  })
}

shinyApp(ui = ui, server = server)
```

We have looked at the median household income (Census) and the percentof adults who are classified as obese (CDC)...now to answer the question we presented at the start of the presentation, we need to show the relationship between household income and obesity. 

```{r}
joinc<-sqldf("select
             a.year, 
             a.state,
             a.income,
             b.state_value
             from joina a
             join joinb b 
             on (a.state=b.state and a.year=b.year)")

nationals<-sqldf("select
year, national_avg_income, national_value
from
(
            select
             a.year, 
             a.national_avg_income,
             b.national_value
             from joina a
             join joinb b 
             on (a.year=b.year)
            )
            group by year, national_avg_income, national_value")


```


##Shiny Q2
```{r}
ui <- fluidPage(
  sidebarPanel(
    selectInput(inputId = "State", label = "State Selected:",
                choices = levels(as.factor(joinc$state))
    ),
  
    helpText(""),
    width = "auto"
  ),
  plotOutput("plot4")
)


server<- function(input, output) {
  
  output$plot4 <-renderPlot({
    

    SelectedState <- input$state
    
    
    ggplot(joinc[joinc$state == SelectedState,]) +
      geom_line(aes(x = state_value, y = income), col = "red", lwd = 1) +
      labs(x="Obesity Rate", y = "Median HHI", 
           title = "Trend of Obesity Rate vs Median HHI Over Time")+
           facet_wrap( ~ joinc$year ) +
      scale_linetype(name = NULL) +
      theme_bw()
  })
}
shinyApp(ui = ui, server = server)
```


```{r}


ggplot(joinc) +
      geom_line(aes(x = state_value, y = income), col = "red", lwd = 1) +
      labs(x="Obesity Rate", y = "Median HHI", 
           title = "Trend of Obesity Rate vs Median HHI Over Time")+
           facet_wrap( ~ joinc$year ) +
      scale_linetype(name = NULL) +
      theme_bw()
```

```{r}
ui <- fluidPage(

    headerPanel("Diamonds Explorer"),
    sidebarPanel(
     selectInput(inputId = "State", label = "State Selected:",
                choices = levels(as.factor(joinc$state))
    )),
    mainPanel(
      plotlyOutput('trendPlot', height = "900px")
    )
)

server <- function(input, output) {

  #add reactive data information. Dataset = built in joinc
  dataset <- reactive(input$state)
  

  output$trendPlot <- renderPlotly({

    # build graph with ggplot syntax
    p <- ggplot(dataset(), aes_string(x = input$state_value, y = input$income)) + 
      geom_point()

    # if at least one facet column/row is specified, add it
    facets <- paste(input$year)
    if (facets != '. ~ .') p <- p + facet_grid(facets)

    ggplotly(p) %>% 
      layout(height = input$plotHeight, autosize=TRUE)

  })

}

shinyApp(ui, server)
```


