---
title: "DATA 608 Week 3"
author: "Vinicio Haro"
date: "March 8, 2019"
output: html_document
runtime: shiny
---

```{r message=FALSE, warning=FALSE}
library(shiny)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
library(DataExplorer)

cdc_url <- "https://raw.githubusercontent.com/charleyferrari/CUNY_DATA608/master/lecture3/data/cleaned-cdc-mortality-1999-2010-2.csv"


cdc <- read.csv(cdc_url, header= TRUE, stringsAsFactors=TRUE)

#change population back into a numeric
cdc$Population <- as.numeric(cdc$Population) 

```


## Simple EDA
We will only get summary statistics for Deaths, Populations, and Crude.Rate since the other fields are strings. We will also look at a missing value plot to check the data integrity. While we are at it, lets take a look at the density distributions for numerical fields to get a sense of the distribution for these fields. 
```{r message=FALSE, warning=FALSE}

head(cdc);
summary(cdc);
plot_missing(cdc);
plot_density(cdc)


```

Lets try to understand the variables at hand before we answer any questions. We are looking at the number of deaths relative to population by year and State. We are also given the crude rate which is the number of new deaths occurring in each state per year, 

## Question 1

As a researcher, you frequently compare mortality rates from particular causes across different States. You need a visualization that will let you see (for 2010 only) the crude mortality rate, across all States, from one cause (for example, Neoplasms, which are effectively cancers). Create a visualization that allows you to rank States by crude mortality for each cause of death.
https://surveillance.cancer.gov/joinpoint/crude.html


```{r message=FALSE, echo=FALSE, fig.height=12, fig.width=12, warning=FALSE}
library(sqldf)
library(ggplot2)
library(rsconnect)

cdc1<-sqldf("
select 
icd_chapter, State, crude_rate 
from 
(
select 
            `ICD.Chapter` as icd_chapter, State, Deaths,Population,power(10,5)*(Deaths/Population) as crude_rate
            from cdc
            where year in ('2010')
            group by `ICD.Chapter`, State
) group by 1,2, 3")





ggplot(cdc1, aes(fill=icd_chapter, y=crude_rate, x=State)) +
    labs(x="State", y="Crude Death Rate per 100,000 Persons") +  
    geom_bar(stat="identity") +
    coord_flip()

```

The static plot has all the information we need, however it is difficult to read since there are numerous icd chapter categories per state. If we deploy this to a shiny app, the user would have more control when it comes to only looking at icd chapters of interest for specific states of interest. We want these elements to be selection category.  
https://shiny.rstudio.com/tutorial/written-tutorial/lesson1/

##Shiny Q1
```{r}
ui <- fluidPage(
  sidebarPanel(
    selectInput(inputId = "CauseDeath", label = "Cause of Death:",
                choices = levels(as.factor(cdc1$icd_chapter))
    ),
    helpText("States by crude mortality for each cause of death."),
    width = "auto"
  ),
  plotOutput("plot1")
)

server<- function(input, output) {
  
  output$plot1 <-renderPlot({
    
    SelectedCause <- input$CauseDeath
    
    ggplot(data=cdc1[cdc1$icd_chapter == SelectedCause,]
           , aes(x = State, y = crude_rate)) +
      labs(x="State", y="Crude Death Rate", 
           title = "Crude death rate per 100,000 persons",
           subtitle = paste("Caused by", SelectedCause)) +  
      geom_bar(stat="identity", fill="steelblue") + 
      geom_hline(aes(yintercept = mean(crude_rate, na.rm = TRUE), linetype = "National Average"), col="red", lwd=1) +
      scale_linetype(name = NULL) +
      theme_bw()
    
  })
}



shinyApp(ui = ui, server = server)

#deployApp()

```




## Question 2
Question 2:
Often you are asked whether particular States are improving their mortality rates (per cause) faster than, or slower than, the national average. Create a visualization that lets your clients see this for themselves for one cause of death at the time. Keep in mind that the national average should be weighted by the national population.

In the previous question, we prepared the data using sql, however it will be easier to use tidyverse methods in order to avoid writing out joins in sql syntax. I will use sqldf to rename the columns. 
```{r message=FALSE, warning=FALSE, fig.height=12, fig.width=12, warning=FALSE}
cdc2 <- cdc %>%
  group_by(Year, ICD.Chapter) %>%
  mutate(N.Population = sum(Population),
         N.Count = sum(Deaths), 
         N.Crude.Rate = 10^5*(N.Count/N.Population)) %>% 
  group_by(Year, ICD.Chapter, State) %>%
  mutate(S.Count=sum(Deaths),
         S.Crude.Rate=10^5*(S.Count/Population)) %>%
  select(ICD.Chapter, State, Year, N.Crude.Rate, S.Crude.Rate)

cdc2b<-sqldf("select 
             'ICD.Chapter' as icd_chapter, 
             State as state,
             Year as year,
             'N.Crude.Rate' as national_crude_rate, 
             'S.Crude.Rate' as state_crude_rate
             from cdc2")
head(cdc2b,10);

ggplot(data=cdc2b, aes(x=year, y=state_crude_rate)) +
      labs(x="Year", y="Crude Death Rate per 100,000 Person") +  
      geom_bar(stat="identity", fill="steelblue") +
      geom_line(aes(x=year, y=national_crude_rate), col="red", lwd=1)
```

The obvious problems with the static graph are the scale and the fact that we can't see the granularity of interest. This is another advantage of the shiny app. The user will have the ability to do apple to apple comparisons for the state and icd chapter of interest. 

##Shiny Q2
```{r}
ui <- fluidPage(
  sidebarPanel(
    selectInput(inputId = "CauseDeath", label = "Cause of Death:",
                choices = levels(as.factor(cdc2$ICD.Chapter))
    ),
    selectInput(inputId = "State", label = "State selected:",
                choices = levels(as.factor(cdc2$State))
    ),
    helpText("Annual Crude mortality rates at State level."),
    width = "auto"
  ),
  plotOutput("plot2")
)

server<- function(input, output) {
  
  output$plot2 <-renderPlot({
    
    SelectedCause <- input$CauseDeath
    SelectedState <- input$State
    
    ggplot(cdc2[cdc2$State == SelectedState & cdc2$ICD.Chapter == SelectedCause,]) +
      geom_bar(aes(x = Year, weight = S.Crude.Rate), fill = "steelblue") +
      labs(x="State", y = "Crude Death Rate", 
           title = "Crude death rate per 100,000 persons",
           subtitle = paste("Caused by", SelectedCause, "in", SelectedState)) +
      geom_line(aes(x = Year, y = N.Crude.Rate, linetype = "National Average"), col = "red", lwd = 1) +
      scale_linetype(name = NULL) +
      theme_bw()
  })
}

shinyApp(ui = ui, server = server)
```

App deployed:
https://vharo00.shinyapps.io/DATA608week3V2/

